### 2019-10-31 LCAforActiveSonarImaging/document.tex
2019-08-18
2019-09-01
2019-10-22
2019-10-30

% !TeX root = III_lca
\title{Low Complexity Adaptive Sonar Imaging}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%
\begin{abstract}
% 
% 
We have studied the low complexity adaptive (LCA) beamformer in active sonar imaging. LCA can be viewed as either a simplification of the minimum variance distortionless response (MVDR) beamformer, or as an adaptive extension to the delay and sum (DAS) beamformer. While both LCA and MVDR attempt to minimize the power of noise and interference in the image, MVDR achieves this by computing optimal array weights from the spatial statistics of the wavefield, while LCA selects the best performing weights out of a predefined set.

@@

MVDR seeks to minimize the power of noise and interference in the output of the beamformer, under the constraint of unity gain in some desired direction $\phi$~\cite{Capon1969}. With the assumption of zero mean data this takes the form of minimizing output variance: 
%
\begin{align}
\argmin{\w[n]}\, E\{\big|z[n]\big|^2\} &= \argmin{\w[n]}\, \w[n]\R[n]\w\H[n]\nn
\text{subject to } \w^H[n]\a_\phi &= 1.\label{III_eq:mvdr_definition}
\end{align}
%
Here $\a_\phi$ is a steering vector for an azimuthal steering angle $\phi$, $E\{\cdot\}$ is the expectation operator, and $\R[n] = E\{\x[n]\x\H[n]\}$ is the spatial covariance matrix for the full array. This is a convex optimization problem with the unique solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a_\phi}{\a_\phi\T\Ri[n]\a_\phi}.\label{III_eq:mvdr_weights}
\end{gather}
The problem lies in estimating and inverting the spatial covariance matrix. We have described the exact steps in~\cite{Buskenes2014}: To avoid signal cancellation we apply spatial averaging by computing a mean covariance matrix from a set of subarrays with length $L$~\cite{Kailath1985}, for true speckle statistics we perform temporal averaging over $N_k = 2K+1$ temporal samples~\cite{Synnevag2009a}, and to improve robustness to parameter errors we add {\large$\epsilon$} percent of the total output power to the matrix's diagonal~\cite{Cox1987,Maksym1979}. These steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

@@

In the image reconstruction we have run MVDR with subarray length $L=16$ and $\text{\large$\epsilon$}=1\%$ diagonal loading. Both MVDR and LCA was run with $K=1$ temporal averaging. This is a fairly aggressive yet stable set of parameters~\cite{Synnevag2009a}.

@@

\ifPhdDoc
\clearpage
\appendix
\renewcommand\thesection{\Alph{section}}
\renewcommand\thesubsection{\Alph{section}.\arabic{subsection}}
\else
\appendices
\fi

\section{Steering angle}\label{III_appendix:steering_angles}

@@

\ifPhdDoc
\typeout{heyheyhey}
   \printbibliography[title=References,heading=subbibliography]
   \addcontentsline{toc}{section}{References}
%    \bibliographysty
%    \bibliography{library.bib}
%    \print
\else
\typeout{hohoho}
   \ifBuildBibliography
      \bibliographystyle{IEEEtran}
      \bibliography{references}

@@

      \bibitem{Synnevag2009a}
      J.-F. Synnev\aa{}g, A.~Austeng, and S.~Holm, ``{Benefits of minimum-variance
      beamforming in medical ultrasound imaging},'' \emph{IEEE Transactions on
      Ultrasonics, Ferroelectrics and Frequency Control}, vol.~56, no.~9, pp.
      1868--1879, Sep. 2009.

>>>>

% !TeX root = III_lca
\title{Low-Complexity Adaptive Sonar Imaging}

@@

% for Computer Society papers, we must declare the abstract and index terms
% PRIOR to the title within the \IEEEcompsoctitleabstractindextext IEEEtran
% command as these need to go into the title area created by \maketitle.
\IEEEcompsoctitleabstractindextext{%

\begin{abstract}
% 
% 
We have studied the low complexity adaptive (LCA) beamformer in active sonar imaging. LCA can be viewed as either a simplification of the minimum variance distortionless response (MVDR) beamformer, or as an adaptive extension to the delay and sum (DAS) beamformer. While both LCA and MVDR attempt to minimize the power of noise and interference in the image, MVDR achieves this by computing optimal array weights from the spatial statistics of the wavefield, while LCA selects the best performing weights out of a predefined set.

@@

MVDR seeks to minimize the power of noise and interference in the output of the beamformer, under the constraint of unity gain in some desired direction $\phi$~\cite{Capon1969}. With the assumption of zero mean data this takes the form of minimizing output variance: 
%
\begin{align}
\argmin{\w[n]}\, E\{\big|z[n]\big|^2\} &= \argmin{\w[n]}\, \w[n]\R[n]\w\H[n]\nn
\text{subject to } \w^H[n]\a_\phi &= 1.\label{III_eq:mvdr_definition}
\end{align}
%
Here $\a_\phi$ is a steering vector for an azimuthal steering angle $\phi$, $E\{\cdot\}$ is the expectation operator, and $\R[n] = E\{\x[n]\x\H[n]\}$ is the spatial covariance matrix for the full array. This is a convex optimization problem with the unique solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a_\phi}{\a_\phi\T\Ri[n]\a_\phi}.\label{III_eq:mvdr_weights}
\end{gather}
The problem lies in estimating and inverting the spatial covariance matrix. We have described the exact steps in~\cite{Buskenes2014}: To avoid signal cancellation we apply spatial averaging by computing a mean covariance matrix from a set of subarrays with length $L$~\cite{Kailath1985}, for true speckle statistics we perform temporal averaging over $N_k = 2K+1$ temporal samples~\cite{Synnevag2009}, and to improve robustness to parameter errors we add {\large$\epsilon$} percent of the total output power to the matrix's diagonal~\cite{Cox1987,Maksym1979}. These steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

@@

In the image reconstruction we have run MVDR with subarray length $L=16$ and $\text{\large$\epsilon$}=1\%$ diagonal loading. Both MVDR and LCA was run with $K=1$ temporal averaging. This is a fairly aggressive yet stable set of parameters~\cite{Synnevag2009}.

@@

\ifPhdDoc
\clearpage
% \renewcommand{\theHchapter}{A\arabic{chapter}}
\renewcommand{\theHsection}{LcaA\arabic{section}}
\appendix
\let\oldthesection\thesection
\let\oldthesubsection\thesubsection
\renewcommand\thesection{\Alph{section}}
\renewcommand\thesubsection{\Alph{section}.\arabic{subsection}}
\else
\appendices
\fi

\section{Steering angle}\label{III_appendix:steering_angles}
% \addcontentsline{toc}{section}{Steering angle}
% \stepcounter{section}

@@

\ifPhdDoc
\let\thesection\oldthesection
\let\thesubsection\oldthesubsection
\typeout{heyheyhey}
   \printbibliography[title=References,heading=subbibliography]
   \addcontentsline{toc}{section}{References}
%    \bibliographysty
%    \bibliography{library.bib}
%    \print
\else
\typeout{hohoho}
   \ifBuildBibliography
      \bibliographystyle{IEEEtran}
      \bibliography{references}

@@

      \bibitem{Synnevag2009}
      J.-F. Synnev\aa{}g, A.~Austeng, and S.~Holm, ``{Benefits of minimum-variance
      beamforming in medical ultrasound imaging},'' \emph{IEEE Transactions on
      Ultrasonics, Ferroelectrics and Frequency Control}, vol.~56, no.~9, pp.
      1868--1879, Sep. 2009.

### 2019-10-31 LCAforActiveSonarImaging/lca (MP078's conflicted copy 2016-05-03).tex
2019-10-30

MVDR seeks to minimize the power of noise and interference in the output of the beamformer, under the constraint of unity gain in some desired direction $\phi$~\cite{Capon1969}. With the assumption of zero mean data this takes the form of minimizing output variance: 
%
\begin{align}
\underset{\w[n]}{\argmin}\, E\{\big|z[n]\big|^2\} &= \underset{\w[n]}{\argmin}\, \w[n]\R[n]\w\H[n]\nn
\text{subject to } \w^H[n]\a_\phi &= 1.\label{eq:mvdr_definition}
\end{align}
%
Here $\a_\phi$ is a steering vector for an azimuthal steering angle $\phi$, $E\{\cdot\}$ is the expectation operator, and $\R[n] = E\{\x[n]\x\H[n]\}$ is the spatial covariance matrix for the full array. This is a convex optimization problem with the unique solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a_\phi}{\a_\phi\T\Ri[n]\a_\phi}.\label{eq:mvdr_weights}
\end{gather}
The problem lies in estimating and inverting the spatial covariance matrix. We have described the exact steps in~\cite{Buskenes2014}: To avoid signal cancellation we apply spatial averaging by computing a mean covariance matrix from a set of subarrays with length $L$~\cite{Kailath1985}, for true speckle statistics we perform temporal averaging over $N_k = 2K+1$ temporal samples~\cite{Synnevag2009a}, and to improve robustness to parameter errors we add {\large$\epsilon$} percent of the total output power to the matrix's diagonal~\cite{Cox1987,Maksym1979}. These steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

>>>>

MVDR seeks to minimize the power of noise and interference in the output of the beamformer, under the constraint of unity gain in some desired direction $\phi$~\cite{Capon1969}. With the assumption of zero mean data this takes the form of minimizing output variance: 
%
\begin{align}
\underset{\w[n]}{\argmin}\, E\{\big|z[n]\big|^2\} &= \underset{\w[n]}{\argmin}\, \w[n]\R[n]\w\H[n]\nn
\text{subject to } \w^H[n]\a_\phi &= 1.\label{eq:mvdr_definition}
\end{align}
%
Here $\a_\phi$ is a steering vector for an azimuthal steering angle $\phi$, $E\{\cdot\}$ is the expectation operator, and $\R[n] = E\{\x[n]\x\H[n]\}$ is the spatial covariance matrix for the full array. This is a convex optimization problem with the unique solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a_\phi}{\a_\phi\T\Ri[n]\a_\phi}.\label{eq:mvdr_weights}
\end{gather}
The problem lies in estimating and inverting the spatial covariance matrix. We have described the exact steps in~\cite{Buskenes2014}: To avoid signal cancellation we apply spatial averaging by computing a mean covariance matrix from a set of subarrays with length $L$~\cite{Kailath1985}, for true speckle statistics we perform temporal averaging over $N_k = 2K+1$ temporal samples~\cite{Synnevag2009}, and to improve robustness to parameter errors we add {\large$\epsilon$} percent of the total output power to the matrix's diagonal~\cite{Cox1987,Maksym1979}. These steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

### 2019-10-31 LCAforActiveSonarImaging/lca_rev1.tex
2019-10-30

MVDR seeks to minimize the output power of the beamformer, under the constraint of a unity gain in some desired direction $\phi$~\cite{Capon1969}:
%
\begin{align}
\underset{\w[n]}{\argmin}\, E\{\big|z[n]\big|^2\} &= \underset{\w[n]}{\argmin}\, \w[n]\R[n]\w\H[n]\nn
\text{subject to } \w[n]\a_\phi &= 1.\label{eq:mvdr_definition}
\end{align}
%
Here $\a_\phi$ is a steering vector, $E\{\cdot\}$ is the expectation operator, and $\R[n] = E\{\x\x\H\}$ is the spatial covariance matrix. This is a convex optimization problem with the solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a_\phi}{\a_\phi\T\Ri[n]\a_\phi},\label{eq:mvdr_weights}
\end{gather}
where $\R=E\{\x[n]\x\H[n]\}\in\mathbb{C}^{M,M}$ is the spatial covariance matrix for the full array. The problem lies in estimating and inverting this spatial covariance matrix. We have described the exact steps in~\cite{Buskenes2014}: To avoid signal cancellation we apply spatial averaging by computing a mean covariance matrix from a set of subarrays with length $L$~\cite{Kailath1985}, for true speckle statistics we perform temporal averaging over $N_k = 2K+1$ temporal samples~\cite{Synnevag2009a}, and to improve robustness to parameter errors we add $d$ percent diagonal loading~\cite{Cox1987,Maksym1979}. These steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

>>>>

MVDR seeks to minimize the output power of the beamformer, under the constraint of a unity gain in some desired direction $\phi$~\cite{Capon1969}:
%
\begin{align}
\underset{\w[n]}{\argmin}\, E\{\big|z[n]\big|^2\} &= \underset{\w[n]}{\argmin}\, \w[n]\R[n]\w\H[n]\nn
\text{subject to } \w[n]\a_\phi &= 1.\label{eq:mvdr_definition}
\end{align}
%
Here $\a_\phi$ is a steering vector, $E\{\cdot\}$ is the expectation operator, and $\R[n] = E\{\x\x\H\}$ is the spatial covariance matrix. This is a convex optimization problem with the solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a_\phi}{\a_\phi\T\Ri[n]\a_\phi},\label{eq:mvdr_weights}
\end{gather}
where $\R=E\{\x[n]\x\H[n]\}\in\mathbb{C}^{M,M}$ is the spatial covariance matrix for the full array. The problem lies in estimating and inverting this spatial covariance matrix. We have described the exact steps in~\cite{Buskenes2014}: To avoid signal cancellation we apply spatial averaging by computing a mean covariance matrix from a set of subarrays with length $L$~\cite{Kailath1985}, for true speckle statistics we perform temporal averaging over $N_k = 2K+1$ temporal samples~\cite{Synnevag2009}, and to improve robustness to parameter errors we add $d$ percent diagonal loading~\cite{Cox1987,Maksym1979}. These steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

### 2019-10-31 LCAforActiveSonarImaging/lca_review1.tex
2019-10-30

MVDR seeks to minimize the output variance of the beamformer, under the constraint of a unity gain in some desired direction $\phi$~\cite{Capon1969}. With the assumption of zero mean data this can be formulated as: 
%
\begin{align}
\underset{\w[n]}{\argmin}\, E\{\big|z[n]\big|^2\} &= \underset{\w[n]}{\argmin}\, \w[n]\R[n]\w\H[n]\nn
\text{subject to } \w^H[n]\a_\phi &= 1.\label{eq:mvdr_definition}
\end{align}
%
Here $\a_\phi$ is a steering vector, $E\{\cdot\}$ is the expectation operator, and $\R[n] = E\{\x[n]\x\H[n]\}$ is the spatial covariance matrix for the full array. This is a convex optimization problem with the solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a_\phi}{\a_\phi\T\Ri[n]\a_\phi}.\label{eq:mvdr_weights}
\end{gather}
The problem lies in estimating and inverting this spatial covariance matrix. We have described the exact steps in~\cite{Buskenes2014}: To avoid signal cancellation we apply spatial averaging by computing a mean covariance matrix from a set of subarrays with length $L$~\cite{Kailath1985}, for true speckle statistics we perform temporal averaging over $N_k = 2K+1$ temporal samples~\cite{Synnevag2009a}, and to improve robustness to parameter errors we add {\large$\epsilon$} percent diagonal loading~\cite{Cox1987,Maksym1979}. These steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

>>>>

MVDR seeks to minimize the output variance of the beamformer, under the constraint of a unity gain in some desired direction $\phi$~\cite{Capon1969}. With the assumption of zero mean data this can be formulated as: 
%
\begin{align}
\underset{\w[n]}{\argmin}\, E\{\big|z[n]\big|^2\} &= \underset{\w[n]}{\argmin}\, \w[n]\R[n]\w\H[n]\nn
\text{subject to } \w^H[n]\a_\phi &= 1.\label{eq:mvdr_definition}
\end{align}
%
Here $\a_\phi$ is a steering vector, $E\{\cdot\}$ is the expectation operator, and $\R[n] = E\{\x[n]\x\H[n]\}$ is the spatial covariance matrix for the full array. This is a convex optimization problem with the solution:
%
\begin{gather}
\vec w[n] = \frac{\Ri[n]\a_\phi}{\a_\phi\T\Ri[n]\a_\phi}.\label{eq:mvdr_weights}
\end{gather}
The problem lies in estimating and inverting this spatial covariance matrix. We have described the exact steps in~\cite{Buskenes2014}: To avoid signal cancellation we apply spatial averaging by computing a mean covariance matrix from a set of subarrays with length $L$~\cite{Kailath1985}, for true speckle statistics we perform temporal averaging over $N_k = 2K+1$ temporal samples~\cite{Synnevag2009}, and to improve robustness to parameter errors we add {\large$\epsilon$} percent diagonal loading~\cite{Cox1987,Maksym1979}. These steps are also needed to ensure that the covariance matrix is numerically well conditioned and hence invertible.

### 2019-10-31 ThesisJP/Paper3/moving_obj.tex
2019-10-30

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). Dotted vertical lines indicate position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996a} using a 96 element phased array with a 22 mm azimuth aperture, which was excited with a 2.5 MHz and 1.5 cycles long pulse. Transmit focus in azimuth and elevation were configured to be equal and the point scatterer was moved through this point. The transmit focus is where the ultrasound beam is at its narrowest; hence shift variance effects will appear at this location first. The beam density was calculated  based on (\ref{eq:resolution}) with a speed of sound equal to 1540. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. %In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located in the center of this beam at one time instance.

>>>>

Figure \ref{fig:das_shiftinvariant} presents such a LSV-plot for an imaging system using the DAS beamformer with a beam spacing equal to (\ref{eq:resolution}). Dotted vertical lines indicate position of transmit-receive beam pairs. The point scatterer response has been simulated using Field II \cite{Jensen1992, Jensen1996} using a 96 element phased array with a 22 mm azimuth aperture, which was excited with a 2.5 MHz and 1.5 cycles long pulse. Transmit focus in azimuth and elevation were configured to be equal and the point scatterer was moved through this point. The transmit focus is where the ultrasound beam is at its narrowest; hence shift variance effects will appear at this location first. The beam density was calculated  based on (\ref{eq:resolution}) with a speed of sound equal to 1540. We see that the contours in Fig.\,\ref{fig:das_shiftinvariant} are approximately diagonal (low P-MAE) and that the amplitude variation among the maximum points is minimal (low A-MAE). We therefore conclude that the system is locally laterally shift-invariant. %In this, and the following LSV-plots based on simulations, the scan sequence always has a beam at zero degrees and the point scatterer is located in the center of this beam at one time instance.

### 2019-10-31 ThesisJo/2_background.tex
2019-10-30

@@

Element-space MVDR often improves image contrast and resolution compared to conventional beamformers~\cite{Benitz1997,Synnevag2007,Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}, but is computationally complex and statistically fragile. It seeks the weight set that minimizes noise and interference accumulated by the array, while enforcing unit gain for signals from the pixel of interest,
%
\begin{gather}
\argmin{\dvec{w}}\ E\Big\{\big|\dvec{w}^H[n]\dvec{x}[n]\big|^2\Big\} \qquad\text{subject to}\qquad \dvec{w}^H[n]\dvec{a} = 1, \label{2_eq_mvdr_def}
% \vec w[n] = \frac{\Ri[n]\1}{\1^T\Ri[n]\1},\label{2_eq_weights}
\end{gather}
%
where $\dvec{a}$ is steering vector. Applying Lagrange multipliers to this convex optimization problem yields the solution
%
\begin{gather}
\dvec w[n] = \frac{\Ri[n]\dvec{a}}{\dvec{a}^T\Ri[n]\dvec{a}} \qquad\text{where}\qquad \dvec{a}=\dvec{1}, \label{2_eq_weights}
\end{gather}
%
where $\dvec{1}$ is a row-vector of ones representing broadside phase-steering and $\R=E\{\dvec{x}[n]\dvec{x}^H[n]\}$ is the spatial covariance matrix for the full array. (\ref{2_eq_weights}) is the maximum likelihood estimate of the signal in the presence of zero-mean random Gaussian noise. We estimate $\R$ with a sample covariance matrix $\eR$, requiring some degree of
%
\begin{itemize}
\item \emph{spatial averaging} to avoid signal cancellation by decorrelating coherent echoes~\cite{Widrow1982,Kailath1985};
\item \emph{temporal averaging} over an interval comparable to the pulse length (one to five samples) to maintain true speckle statistics~\cite{Synnevag2009a};
\item \emph{diagonal loading} to improve robustness to parameter errors~\cite{Cox1987,Maksym1979,VanTrees2002}.
\end{itemize}%

requiring some degree of \emph{spatial averaging} to avoid signal cancellation, \emph{temporal averaging} to maintain true speckle statistics, and \emph{diagonal loading} to improve robustness to parameter errors~\cite{Synnevag2009a}.

Combined, these techniques also ensure a numerically well conditioned $\eR$.

\begin{itemize}
\item \emph{spatial averaging} to avoid signal cancellation by decorrelating coherent echoes~\cite{Kailath1985};
\item \emph{temporal averaging} over an interval comparable to the pulse length (one to five samples) to maintain true speckle statistics~\cite{Synnevag2009a};
\item \emph{diagonal loading} to improve robustness to parameter errors~\cite{Cox1987,Maksym1979}.
\end{itemize}%

We split the array into subarrays of length $L$ that overlap on all but one sensor. If we let $x_l[n]$ represent the datavector from subarray $l$,
%
\begin{gather}
\dvec{x}_l[n] = \bmat{x_l[n] & x_{l+1}[n] & \dots & x_{l+L-1}[n]}^T,
\end{gather}
%
then the sample covariance matrix formed with temporal and spatial averaging, $\eR$, is given as
%
\begin{gather}
\eR[n] =  \frac{1}{N_K N_L} \sumb{l=0}{N_L-1}\sumb{n'=n-K}{n+K} \x_l[n]\x_l^H[n] \in\mathbb{C}^{L,L},\label{2_eq_spatialR}
\end{gather}
%
where $N_K = 2K+1$ is the number of temporal samples to perform averaging over, and $N_L = M-L+1$ is the number of subarrays. In the actual design a fraction of the total power of $\eR[n]$ is finally added to its diagonal~\cite{Synnevag2007}, but this step will be ignored here due to its negligible contribution to runtime.

@@

Above 32 channels the inversion step becomes a greater computational bottleneck than covariance estimatation. We mitigate this by applying space reduction in the spatial frequency-domain, or beamspace. If the data is narrow band, narrow beam, or both, the backscatter energy resides primarily in the center frequencies. The zero-frequency correspond to broadside, higher frequencies to larger opening angles. Since the combined transmit and receive response is highly directive, it is sufficient to process only three or five center angles and ignore the rest~\cite{Nilsen2009a,VanTrees2002}. The method is called beamspace MVDR.

Beamspace MVDR formulation closely resembles that of time domain MVDR, the main difference being an initial need to Fourier transform the data:
%
\begin{gather}
\x_{\text{BS},l}[n] = \mat{B}\x_l[n], \qquad \mat{B}\in \mathbb{C}^{N_b,L}, \qquad \text{where} \qquad [\mat{B}]_{p,q} = \frac{1}{\sqrt{L}} e^{-j2\pi\,p\,q/L}.\label{2_eq_beamspace}
\end{gather}
%
Here $\mat{B}$ refers to the Butler matrix, $l$ is subarray index, and $N_b$ is number of beams that should be computed. Note has once $\x_{\text{BS},l}[n]$ is computed, $\eR_\text{BS}$ can be found as in (\ref{2_eq_spatialR}), $\w_\text{BS}$ as in (\ref{2_eq_weights}), and $z$ as in (\ref{z}).

Reducing the covariance size from $L$ to $N_b$ also reduce the estimation step to O($N_K N_L N_b^2$), and the inversion step to O($N_b^3$). The downside is the additional computations required by the frequency transform, which must be weighted up against the gain of reducing $N_b$.

>>>>

% MVDR

% MPDR

% Applications: Radar, radar astronomy, sonar, commnications, direction finding, seismology, tomography,

% Performance measures
% Array: Directivity, array gain vs spatially white noise, sensitivity or tolerance factor,

@@

Element-space (time-domain) MVDR often improves image contrast and resolution compared to conventional beamformers~\cite{Benitz1997,Synnevag2007,Blomberg2013,Blomberg2012a,Dursun2009,Lo2004}, but is computationally complex and statistically fragile (more on this later). It seeks the weight set that minimizes noise and interference accumulated by the array, while enforcing unit gain for signals from the pixel of interest,
%
\begin{gather}
\argmin{\dvec{w}}\ E\Big\{\big|\dvec{w}^H[n]\dvec{x}[n]\big|^2\Big\} \qquad\text{subject to}\qquad \dvec{w}^H[n]\dvec{a} = 1, \label{2_eq_mvdr_def}
% \vec w[n] = \frac{\Ri[n]\1}{\1^T\Ri[n]\1},\label{2_eq_weights}
\end{gather}
%
where $\dvec{a}$ is steering vector. Solving this convex optimization problem with Lagrange multipliers yields
%
\begin{gather}
\dvec w[n] = \frac{\Ri[n]\dvec{a}}{\dvec{a}^T\Ri[n]\dvec{a}}, \label{2_eq_weights}
\end{gather}
%
where $\dvec{a}=\dvec{1}$ is a row-vector of ones representing broadside phase-steering and $\R=E\{\dvec{x}[n]\dvec{x}^H[n]\}$ is the spatial covariance matrix for the full array. In the presence of zero-mean random Gaussian noise, these weights yield the maximum likelihood estimate of the signal and the maximum signal-to-interference-and-noise ratio (SINR). However, active systems or passive ones subjected to multipath or jamming, experience highly correlated noise, interference and signal~\cite{Widrow1982}. Limited amount of data, array miscalibration and model inaccuracies makes matters worse, sometimes resulting in performance inferior to DAS~\cite{Li2006robust}.

To make MVDR robust, it must handle avoid signal cancellation, preserve speckle statistics and ensure numerical and statistical robustness. The literature suggests estimating $\R$ with a sample covariance matrix $\eR$ using some degree of
%
\begin{itemize}
\item \emph{Spatial smoothing}: Decorrelates coherent echoes to avoid signal cancellation, assuming regular arrays. We perform subarray averaging~\cite{Kailath1985}, which trades off spatial resolution and makes the covariance matrix exhibit Toeplitz-like structure. Some studies construct such matrices directly since are simpler to invert [TK87]. $\eR$ is guaranteed non-singular in cases with twice as many subarrays as signals [WPMS88,PK89]. 
\item \emph{Temporal smoothing} over an interval comparable to the pulse length (one to five samples) to maintain true speckle statistics~\cite{Synnevag2009};
\item \emph{Diagonal loading} to improve robustness to parameter errors~\cite{Cox1987,Maksym1979,VanTrees2002}.
\end{itemize}%
%
These techniques ensure a numerically well conditioned $\eR$, allow reducing the computational complexity and expose enough data parallelism to facilitate GPU-acceleration.

Topeplitz [TK87] - choose wm to make the smoothed matrix close to Toeplitz. Topelitz weighting, adaptive spatial averaging in literature. [TK87]

Formulation of the sample covariance matrix $\eR$ begins by splitting the array into subarrays of length $L$ that overlap on all but one sensor. If we let $x_l[n]$ represent the datavector from subarray $l$,
%
\begin{gather}
\dvec{x}_l[n] = \bmat{x_l[n] & x_{l+1}[n] & \dots & x_{l+L-1}[n]}^T,
\end{gather}
%
then $\eR$ formed with temporal and spatial averaging (smoothing) is given as
%
\begin{gather}
\eR[n] =  \frac{1}{N_K N_L} \sumb{l=0}{N_L-1}\sumb{n'=n-K}{n+K} \x_l[n]\x_l^H[n] \in\mathbb{C}^{L,L},\label{2_eq_spatialR}
\end{gather}
%
where $N_K = 2K+1$ is the number of temporal samples to perform averaging over, and $N_L = M-L+1$ is the number of subarrays. A fraction of the total power of $\eR[n]$ is finally added to its diagonal~\cite{Synnevag2007}.

@@

Systems with narrow beams, narrow bands, or both, have most backscattered energy residing in center spatial frequencies. Zero-frequency maps to broadside, higher frequencies map to larger opening angles. Computing weighted combinations of such frequencies, similarly to doing so with elements, is referred to as beamspace or discrete Fourier-transform (DFT) beamforming.

%,  Similarly to performing weighted sums of elements, we perform weighted sums of frequencies  Weighted sums of Performing weighted sum

In imaging sonar and ultrasound systems, the combined transmit and receive response is usually narrow enough to contain most energy in three or five center beams~\cite{Nilsen2009a,VanTrees2002}. This allows  Above 32 channels the inversion step becomes a greater computational bottleneck than covariance estimation. We mitigate this by applying space reduction in the spatial frequency-domain, or beamspace.

Beam-space MVDR resembles element-space MVDR, but demands an initial Fourier transformation of the data:
%
\begin{gather}
\x_{\text{BS},l}[n] = \mat{B}\x_l[n], \qquad \mat{B}\in \mathbb{C}^{N_b,L},\label{2_eq_beamspace}
\end{gather}
%
where $l$ is subarray index, $N_b$ is number of beams that should be computed and
%
\begin{gather}
[\mat{B}]_{p,q} = \frac{1}{\sqrt{L}} e^{-j2\pi\,p\,q/L}.\label{2_eq_butler}
\end{gather}
%
is the Butler (or DFT) matrix~\cite{Butler1961}. Once $\x_{\text{BS},l}[n]$ is computed, $\eR_\text{BS}$ can be found as in (\ref{2_eq_spatialR}), $\w_\text{BS}$ as in (\ref{2_eq_weights}), and $z$ as in (\ref{z}).

Reducing $\eR_\text{BS}$ size from $L$ to $N_b$ reduce estimation to O($N_K N_L N_b^2$), and the inversion step to O($N_b^3$). The downside is the additional computations required by the frequency transform, which must be weighted up against the gain of reducing $N_b$.
